---
title: 'Reading In Data'
date: Last updated \today
author:
- name: Shiro Kuriwaki
output:
  pdf_document:
    latex_engine: xelatex
    template: xelatex.template
    fig_caption: TRUE
---


```{r, include = FALSE}
library(tidyverse)
library(survey)
library(srvyr)
knitr::opts_chunk$set(echo = TRUE)
```

By now, you should have:

* Downloaded R or Stata
* Understood wht's what in a GUI

# Where is Your File?

Computer files (data, documents, programs) are organized hiearchically, like a branching tree. Folders can contain files, and also other folders. To load a dataset, you need to specify where that file is. 

We denote the hierarchy of a folder by the `/` (slash) symbol. `data/input/2018-08` indicates the `2018-08` folder, which is included in the `input` folder, which is in turn included in the `data` folder. 

Files (but not folders) have "file extensions" which you are probably familiar with already: `.docx`, `.pdf`, and `.pdf` already. The file extensions you will see for this course a lot are:

* `.csv`: A comma separated values file
* `.xlsx`
* `.dta`: Stata data
* `.sav`: SPSS data
* `.Rds`: R data single file
* `.Rdata`: R data (could be multiple files)



The goal of the rest of this document is to enable you to write code that will read in data files into your GUI. 

# Reading in R

Use the `haven` package. There is a function to read and write certain types of data files. 
```{r}
library(haven)
```


```{r}
hh <- readRDS("data/input/2018-08/HHP_August2018_data.Rds")
```

You can call the name you gave it to browse it:

```{r, eval = FALSE}
hh
```


For a discrete outcome like a poll question, the `count` function is quick and useful. Enter the name of your dataset and the name of the variable you care about. 

To use the `count` function you need the `dplyr` package. Add `library(dplyr)` to the beginning of your code. Even easier, add `library(tidyverse)`, that wil load `dplyr` and other related packages we will end up using. 

```{r}
library(tidyverse)
```

Now let's use `count`

```{r}
count(hh, QM3AR1)
```

# Reading in Stata


The command to read in a Stata file is `use`

```{javascript, eval = FALSE}
use HHP_August2018_data.dta, clear
```

Actually, probably the harder part is getting your computer to find that file. In my case, it looks like

```{shell, eval = FALSE}
cd "~/Dropbox/survey-notes/data/input/2018-08"
```


