---
title: 'Tables from Data'
date: Last updated \today
author:
- name: Shiro Kuriwaki
output:
  pdf_document:
    latex_engine: xelatex
    template: xelatex.template
    fig_caption: TRUE
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

By now, you should have:

* Downloaded R or Stata
* Understood wht's what in a GUI
* Read in the August 2018 Poll

# Reading In


As we saw in the previous note, in R:

```{r, message = FALSE}
library(tidyverse)

hh <- readRDS("data/input/2018-08/HHP_August2018_data.Rds")
```



And in Stata reading in a Stata dataset  is through the `use` command

```{javascript, eval = FALSE}
cd "data/input/2018-08"
use HHP_August2018_data, clear
```


# Simple Tables in R


```{r}
xtabs(formula = ~ QHO1, data =  hh)
```

And in R, you can abbreviate the argument names if you give them in the order provided:

```{r}
xtabs(~ QM3AR1 + QHO1, hh)
```

# Simple Tables in Stata

The tab (or tabulate) command
```{javascript, eval = FALSE}
tab QM3AR1
```

```
       M3AR1 1. The |
 economy: M3 Do you |
         approve or |
  disapprove of the |
job President Trump |      Freq.     Percent        Cum.
--------------------+-----------------------------------
   Strongly approve |        414       29.01       29.01
   Somewhat approve |        381       26.70       55.71
Somewhat disapprove |        244       17.10       72.81
Strongly disapprove |        388       27.19      100.00
--------------------+-----------------------------------
              Total |      1,427      100.00

```

Options come after the comma (look at help page for what options are available)

```{javascript, eval = FALSE}
tab QM3AR1, sort
```


```
       M3AR1 1. The |
 economy: M3 Do you |
         approve or |
  disapprove of the |
job President Trump |      Freq.     Percent        Cum.
--------------------+-----------------------------------
   Strongly approve |        414       29.01       29.01
Strongly disapprove |        388       27.19       56.20
   Somewhat approve |        381       26.70       82.90
Somewhat disapprove |        244       17.10      100.00
--------------------+-----------------------------------
              Total |      1,427      100.00
```

Cross-tabs

```{javascript, eval = FALSE}
tab QM3AR1 QHO1 
```

```
       M3AR1 1. The |
 economy: M3 Do you |
         approve or |    HO1 Do you have
  disapprove of the |   health insurance?
job President Trump |       Yes         No |     Total
--------------------+----------------------+----------
   Strongly approve |       375         39 |       414 
   Somewhat approve |       352         29 |       381 
Somewhat disapprove |       220         24 |       244 
Strongly disapprove |       350         38 |       388 
--------------------+----------------------+----------
              Total |     1,297        130 |     1,427 
```



# Defining the Survey in R

To do more complex operations like using weights, use the `survey` package to define the properties of the dataset.

```{r, message = FALSE}
library(survey)
library(srvyr)
```


To set weights, 

```{r}
hh_svy <- svydesign(data = hh, ids = ~0, weights = ~Propwts)
```

To do cross-tabs (now with weights)

```{r}
svytable(~ QM3AR1 + QHO1, hh_svy)
```

